Deep Reinforcement Learning for Conservation Decisions

### 2020-12-12 manuscript/appendix.Rmd
td3 + a2c results & cache

---
title: Appendix for: Deep Reinforcement Learning for Conservation Decisions
authors:
  - name: Marcus Lapeyrolerie
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
  - name: Melissa Chapman
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
  - name: Kari Norman
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
  - name: Carl Boettiger
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: cboettig@berkeley.edu

@@

```{r knitr, include = FALSE}
knitr::opts_chunk$set(echo=FALSE)
```

```{r setup, cache = FALSE}
library(tidyverse)
library(patchwork)
library(reticulate)

@@

# force CPU-evaluation for reproducibility
Sys.setenv("CUDA_VISIBLE_DEVICES"="")

@@

```{r eval=!file.exists("cache/a2c.zip")}
# train an agent (model) on one of the environments:

@@

```{r}
# Simulate management under the trained agent
#td3 <- sb3$A2C$load("cache/td3") ## fixme, fails to load
sims <- env$simulate(td3, reps = 500L)
policy <- env$policyfn(td3, reps = 50L)
```

```{r}
# Simulate under the optimal solution (given the model)
opt <- gym_fishing$models$escapement(env)
opt_sims <- env$simulate(opt, reps = 500L)
opt_policy <- env$policyfn(opt) %>% rename(optimal = action)
```

```{r}
sims_df <- bind_rows(sims, opt_sims, .id = "model") %>%
  mutate(model = c("TD3", "optimal")[as.integer(model)])

policy <- left_join(policy, opt_policy)
```

```{r}
gamma <- 1 #discount
fig_reward <- sims_df %>% 
  group_by(rep, model) %>%
  mutate(cum_reward = cumsum(reward * gamma^time)) %>%
  group_by(time, model) %>%
  summarise(mean_reward = mean(cum_reward), 
            sd = sd(cum_reward), .groups = "drop") %>% 
  ggplot(aes(time, mean_reward)) + 
  geom_ribbon(aes(ymin = mean_reward - 2*sd, 
                  ymax = mean_reward + 2*sd, fill = model),
              alpha=0.25, show.legend = FALSE) + 
  geom_line(aes(col = model)) + ylab("mean cumulative reward")
```

@@

```{r}
fig_policy <- 
  policy %>% ggplot(aes(state, state - action, group=rep)) + 
  geom_point(alpha=.05) + geom_line(aes(y = state - optimal), col = "blue") + 
  coord_cartesian(xlim = c(0, 1.5), ylim=c(0,0.9))
```

>>>>

---
title: "Appendix for: Deep Reinforcement Learning for Conservation Decisions"
authors:
  - name: Marcus Lapeyrolerie
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
  - name: Melissa Chapman
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
  - name: Kari Norman
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
  - name: Carl Boettiger
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: cboettig@berkeley.edu

@@

```{r knitr, include = FALSE}
knitr::opts_chunk$set(echo=FALSE, message = FALSE, warning = FALSE,
                      fig.width = 7, fig.height = 4)
ggplot2::theme_set(ggplot2::theme_bw())

scale_colour_discrete <- function(...) ggthemes::scale_colour_solarized()
scale_fill_discrete <- function(...) ggthemes::scale_fill_solarized()
pal <- ggthemes::solarized_pal()(6)
txtcolor <- "#586e75"
```

```{r setup, cache = FALSE, message=FALSE}
library(tidyverse)
library(patchwork)
library(reticulate)

@@

# force CPU-evaluation if needing perfect reproducibility
Sys.setenv("CUDA_VISIBLE_DEVICES"="0")

@@

```{r eval=!file.exists("cache/td3.zip")}
# train an agent (model) on one of the environments:

@@

```{r}
# Simulate management under the trained agent
td3 <- sb3$TD3$load("cache/td3")
td3_sims <- env$simulate(td3, reps = 500L)
td3_policy <- env$policyfn(td3, reps = 50L)
```

```{r eval=!file.exists("cache/a2c.zip")}
# train an agent (model) on one of the environments:
a2c <- sb3$A2C('MlpPolicy', env, verbose=0L, seed = 42L) # Must use L for integers!
a2c$learn(total_timesteps=300000L)
a2c$save("cache/a2c")
```

```{r}
# Simulate management under the trained agent
a2c <- sb3$A2C$load("cache/a2c")
a2c_sims <- env$simulate(a2c, reps = 500L)
a2c_policy <- env$policyfn(a2c, reps = 50L)
```

```{r}
# Simulate under the optimal solution (given the model)
opt <- gym_fishing$models$escapement(env)
opt_sims <- env$simulate(opt, reps = 500L)
opt_policy <- env$policyfn(opt)
```

```{r}
sims_df <- bind_rows(td3_sims, a2c_sims, opt_sims, .id = "model") %>%
  mutate(model = c("TD3", "A2C", "optimal")[as.integer(model)])

policy_df <- bind_rows(td3_policy, a2c_policy, opt_policy, .id = "model") %>%
  mutate(model = c("TD3", "A2C", "optimal")[as.integer(model)])

gamma <- 1 #discount
reward_df <- sims_df %>% 
  group_by(rep, model) %>%
  mutate(cum_reward = cumsum(reward * gamma^time)) %>%
  group_by(time, model) %>%
  summarise(mean_reward = mean(cum_reward), 
            sd = sd(cum_reward), .groups = "drop")

@@

```{r}
fig_policy <- 
  policy_df %>% ggplot(aes(state, state - action, 
                           group=interaction(rep, model),
                           col = model)) + 
  geom_line(show.legend = FALSE) + 
  coord_cartesian(xlim = c(0, 1.3), ylim=c(0,0.9)) + 
  ylab("escapement")
```

```{r}

fig_reward <- reward_df %>% 
  ggplot(aes(time, mean_reward)) + 
  geom_ribbon(aes(ymin = mean_reward - 2*sd, 
                  ymax = mean_reward + 2*sd, fill = model),
              alpha=0.25, show.legend = FALSE) + 
  geom_line(aes(col = model), show.legend = FALSE) + 
  ylab("reward")
```

### 2020-12-12 manuscript/manuscript.Rmd
td3 + a2c results & cache

```{r setup, cache = FALSE, include = FALSE}
library(tidyverse)
library(patchwork)
library(reticulate)

knitr::opts_chunk$set(echo=FALSE)
```

@@

fishing gym?

```{r cache=TRUE}
## Python dependencies
gym         <- import ("gym")
gym_fishing <- import("gym_fishing")
sb3         <- import ("stable_baselines3")

# train an agent (model) on one of the environments:
env <- gym$make("fishing-v1")
a2c <- sb3$A2C('MlpPolicy', env, verbose=0L, seed = 42L) # Must use L for integers!
a2c$learn(total_timesteps=300000L)
sims <- env$simulate(a2c, reps = 500L)
```

```{r}
fig_fish_sims <- sims %>% 
  filter(rep < 50) %>%
  ggplot(aes(time, state, group = as.factor(rep))) + geom_line(alpha=0.1)
policy <- env$policyfn(a2c, reps = 50L)
fig_fish_policy <- policy %>% ggplot(aes(state, state - action, group=rep)) + geom_point()
```

```{r fig1, fig.cap=""}
fig_fish_sims + fig_fish_policy
```

>>>>

@@

Note that the A2C policy learns to shut down fishing entirely below a certain stock size, without any variation.  However, it keeps the fishery closed even once stock sizes are slightly higher than $B_{MSY}$.  Once it begins fishing, it slightly under-harvests for stock sizes a little over $B_{MSY}$, but tends to over-harvest for very large stock sizes (rarely ever realized, as fishing pressure is enough to keep it out of this range). Note also that the harvest varies across replicates, creating a fuzzy region in the policy function. Under such a policy, A2C maintains a stock size on average significantly above $B_{MSY}$, though also sees higher volatility particularly on the low side.  Note that due to the stochastic nature of the environment, any given replicate A2C may out-perform the optimal policy, though on average the optimal policy provides consistently better yields.
